---
title: "Data606_lab5A: Foundations for statistical inference - Sampling distributions"
author: "Mahmud Hasan Al Raji"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r }
library(tidyverse)
library(openintro)
library(infer)
library(ggplot2)
```

# Creating the Data

The Wellcome Global Monitor finds that 20% of people globally do not believe that the work scientists do benefits people like them. In this lab, I will assume this 20% is a true population proportion and learn about how sample proportions can vary from sample to sample by taking smaller samples from the population. I will first create our population assuming a population size of 100,000. This means 20,000 (20%) of the population think the work scientists do does not benefit them personally and the remaining 80,000 think it does.

```{r }
global_monitor <- tibble(
  scientist_work = c(rep("Benefits", 80000), rep("Doesn't benefit", 20000))
)
```

Here, the name of the data frame is global_monitor and the name of the variable that contains responses to the question “Do you believe that the work scientists do benefit people like you?” is scientist_work.

Quick visualization of the distribution of responses above using a bar plot:

```{r }
ggplot(global_monitor, aes(x = scientist_work)) +
  geom_bar() +
  labs(
    x = "", y = "",
    title = "Do you believe that the work scientists do benefit people like you?"
  ) +
  coord_flip() 
```
obtaining summary statistics to confirm that the data frame is constrcuted correctly.

```{r }
global_monitor %>%
  count(scientist_work) %>%
  mutate(p = n /sum(n))
```

# The unknown sampling distribution

collecting a simple random sample of size 50 from the global_monitor dataset, and assigning the result to samp1

```{r }
set.seed(03021)
samp1 <- global_monitor %>%
  sample_n(50)
```

# Exercise1: Describe the distribution of responses in this sample. How does it compare to the distribution of responses in the population. 

```{r }
samp1 %>%
  count(scientist_work) %>%
  mutate(p_hat = n /sum(n))
```

** Ans1: In the above distribution of random sample of 50 people (less than 1% of the population), it is found that the proportion of people believe that sicentist works don't benefit them is 0.26. That means it is a bit above the true population proportion of 0.20 in the similar belief context.

# Exercise2: Would you expect the sample proportion to match the sample proportion of another student’s sample? Why, or why not? If the answer is no, would you expect the proportions to be somewhat different or very different? Ask a student team to confirm your answer.

** Ans2: I dont expect that this sample porportion will be exactly matched with another student's sample proportion. Because every sample here is being randomly selected with different values. If the sample size be remained same but with different values the proprtions will be somewhat different.

# Exercise3: Take a second sample, also of size 50, and call it samp2. How does the sample proportion of samp2 compare with that of samp1? Suppose we took two more samples, one of size 100 and one of size 1000. Which would you think would provide a more accurate estimate of the population proportion?

```{r }
set.seed(0409)
samp2 <- global_monitor %>%
  sample_n(50)

samp2 %>%
  count(scientist_work) %>%
  mutate(p_hat = n /sum(n))
```

** Ans3: The sample proportion of samp2 is little bit lower than the sample proportion of samp1. If I take two more samples of sizes of 100 and 1000, I think the sample of size 1000 will provide a more accurate estimate of the population proportion. As per the Central Limit Theorem,the larger sample size tends to provide more accurate estimate of the population proportion.

# Building up the sampling distribution for the sample proportion: Here, I have taken 15,000 different samples of size 50 from the population, calculated the proportion of responses in each sample, filter for only the Doesn’t benefit responses, and stored each result in a vector called sample_props50. Here, I have specified that replace = TRUE since sampling distributions are being constructed by sampling with replacement.

```{r }
set.seed(3456)
sample_props50 <- global_monitor %>%
                    rep_sample_n(size = 50, reps = 15000, replace = TRUE) %>%
                    count(scientist_work) %>%
                    mutate(p_hat = n /sum(n)) %>%
                    filter(scientist_work == "Doesn't benefit")
```

# visualization of the distribution of the above proportions with a histogram.

```{r }
ggplot(data = sample_props50, aes(x = p_hat)) +
  geom_histogram(binwidth = 0.02) +
  labs(
    x = "p_hat (Doesn't benefit)",
    title = "Sampling distribution of p_hat",
    subtitle = "Sample size = 50, Number of samples = 15000"
  )

```

# Exercise4: How many elements are there in sample_props50? Describe the sampling distribution, and be sure to specifically note its center. Make sure to include a plot of the distribution in your answer.

** Ans: 15000 samples are present in sample_props50. The plot looks mostly normally distributed, with a slight skew to the right.The center seems to at around 0.2 which is consistent with the population distribution.

# Exercise5: To make sure you understand how sampling distributions are built, and exactly what the rep_sample_n function does, try modifying the code to create a sampling distribution of 25 sample proportions from samples of size 10, and put them in a data frame named sample_props_small. Print the output. How many observations are there in this object called sample_props_small? What does each observation represent?


```{r  }
set.seed(8987)
sample_props_small<-global_monitor%>%
          rep_sample_n(size=10, reps=25,replace=TRUE)%>%
          count(scientist_work) %>%
           mutate(p_hat = n /sum(n)) %>%
           filter(scientist_work == "Doesn't benefit")
sample_props_small
```

** Ans5: sample_props_small’ represent a sample distribution of 25 observations. Each observation represents a sample proportion.

# Sample size and the sampling distribution

```{r }
ggplot(data = sample_props50, aes(x = p_hat)) +
  geom_histogram(binwidth = 0.02)
```

# Exercise6: Use the app below to create sampling distributions of proportions of Doesn’t benefit from samples of size 10, 50, and 100. Use 5,000 simulations. What does each observation in the sampling distribution represent? How does the mean, standard error, and shape of the sampling distribution change as the sample size increases? How (if at all) do these values change if you increase the number of simulations? (You do not need to include plots in your answer.)

Ans: Each observation in the sampling distribution represents the mean of each sample. As the sample size increases the mean is getting closer to the true population proportion of 0.20 and the standard error is decreasing because of less variability.Also, the shape of the sampling distribution is becoming normal.The number of simulations have no impact on the mean and standard error.  


# Exercise7: Take a sample of size 15 from the population and calculate the proportion of people in this sample who think the work scientists do enhances their lives. Using this sample, what is your best point estimate of the population proportion of people who think the work scientists do enchances their lives?

```{r }
set.seed(9088)

samp3<-global_monitor%>%
      sample_n(15) %>% 
      count(scientist_work) %>% 
      mutate(p_hat= n/sum(n)) %>% 
     filter(scientist_work=="Benefits")
samp3
```

** Ans7: The best point estimate of the population proportion of people who think the work scientists do enhances their lives is 0.8 i.e 80% people of this smaple size of 15 think that scientists work do enhances their lives.

# Exercise8: Since you have access to the population, simulate the sampling distribution of proportion of those who think the work scientists do enchances their lives for samples of size 15 by taking 2000 samples from the population of size 15 and computing 2000 sample proportions. Store these proportions in as sample_props15. Plot the data, then describe the shape of this sampling distribution. Based on this sampling distribution, what would you guess the true proportion of those who think the work scientists do enchances their lives to be? Finally, calculate and report the population proportion.

```{r }
set.seed(4987)

sample_props15<-global_monitor %>% 
              rep_sample_n(size=15, reps=2000, replace = TRUE) %>% 
              count(scientist_work) %>% 
              mutate(p_hat=n/sum(n)) %>% 
              filter(scientist_work=="Benefits")

ggplot(sample_props15,aes(p_hat))+geom_histogram(binwidth = 0.02) +
  labs(
    x = "p_hat (Benefits)",
    title = "Sampling distribution of p_hat",
    subtitle = "Sample size = 15, Number of samples = 2000"
  )

mean(sample_props15$p_hat)
```
** Ans8: It is seen from the plot above that the shape of the distribution is skewed to the left.The true proportion of those who think the work scientists do enchance their lives is 0.8014.The population proportion is approximately the mean of the sampling distribution which is 0.8014.

# Exercise9: Change your sample size from 15 to 150, then compute the sampling distribution using the same method as above, and store these proportions in a new object called sample_props150. Describe the shape of this sampling distribution and compare it to the sampling distribution for a sample size of 15. Based on this sampling distribution, what would you guess to be the true proportion of those who think the work scientists do enchances their lives?

```{r }
set.seed(74567)

sample_props150<-global_monitor %>% 
              rep_sample_n(size=150, reps=2000, replace = TRUE) %>% 
              count(scientist_work) %>% 
              mutate(p_hat=n/sum(n)) %>% 
              filter(scientist_work=="Benefits")

ggplot(sample_props150,aes(p_hat))+geom_histogram(binwidth = 0.02) +
  labs(
    x = "p_hat (Benefits)",
    title = "Sampling distribution of p_hat",
    subtitle = "Sample size = 150, Number of samples = 2000"
  )

mean(sample_props150$p_hat)

```
** Ans9: Based on the sampling above it is found that the shape of this sampling distribution seems normal. It does not have skewness like the the sampling distribution of sample size of 15. The true proportion of population who think the work of scientists do enhance their lives is the mean of the sampling distribution, which is 0.799.

# Exercise10: Of the sampling distributions from 2 and 3, which has a smaller spread? If you’re concerned with making estimates that are more often close to the true value, would you prefer a sampling distribution with a large or small spread?

** Ans10: The sampling distribution with 150 sample size has smaller spread. It is seen above that the sampling distribution with small spread provides estimate that is more close to the true value, I would prefer a sampling distribution with a small spread.


