---
title: "Data606_lab5B:Foundations for statistical inference - Confidence intervals"
author: "Mahmud Hasan Al Raji"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r }

library(tidyverse)
library(openintro)
library(infer)
library(ggplot2)
```

# Creating data frame based on Pew research assumption that 62% of total 100000 US adults think  climate change impacts their commuinity. The rest percentage of adults think reverse.

```{r }
us_adults <- tibble(
  climate_change_affects = c(rep("Yes", 62000), rep("No", 38000))
)
```

# Visualization of the distribution of these responses using a bar plot

```{r }
ggplot(us_adults, aes(x = climate_change_affects)) +
  geom_bar() +
  labs(
    x = "", y = "",
    title = "Do you think climate change is affecting your local community?"
  ) +
  coord_flip() 
```
# Summary statistics to confirm that the data frame is constructed correctly

```{r }

us_adults %>%
  count(climate_change_affects) %>%
  mutate(p = n /sum(n))

```

# Now consider a sample size of 60. Below is code to see the summary statistics

```{r }
set.seed((1234))
samp <- us_adults %>%
  sample_n(size = 60)
samp %>% count(climate_change_affects) %>% mutate(p_hat=n/sum(n))
```

# Exercise 1: What percent of the adults in your sample think climate change affects their local community?

** Ans1: 61.67% adults of the sample above think that climate change affects their local community.

# Exercise 2 : Would you expect another student’s sample proportion to be identical to yours? Would you expect it to be similar? Why or why not?

** Ans2: I don't think that another student's sample proportion will be identical to my sample proportion. Because each sample values will be diffrent even though the sample size remains same. It is only possible to get the same sample proportion if both the sample's values are exactly same. But that's a rare thing to happen I believe.

# Confidence level

```{r }
set.seed(4567)
samp %>%
  specify(response = climate_change_affects, success = "Yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)

```

# Exercise 3: In the interpretation above, we used the phrase “95% confident”. What does “95% confidence” mean?

** Ans3: 95% confidence means that we are 95% certain or sure of something. Thus,95% confidence interval is a range of values that is 95% certain will contain the true proportion of the population.
 
# Exercise 4: Does your confidence interval capture the true population proportion of US adults who think climate change affects their local community? If you are working on this lab in a classroom, does your neighbor’s interval capture this value?

** Ans4: Yes, my confidence interval capture the true population proportion of US adults who think climate change affects their local community. My confidence interval is, ci = (0.48, 0.73). The true population proportion of US adults who think climate change affects their local community is 0.62, which is within my confidence interval. 

# Exercise 5: Each student should have gotten a slightly different confidence interval. What proportion of those intervals would you expect to capture the true population mean? Why?

** Ans5: Each student should get a slightly different confidence interval due to difference in different sample values. As, all the students have worked with 95% confidenece interval level it is 95% certian for all that the true population proportion is contained in the stated confidence interval. So, I would expect at least 95% of those intervals to capture the true population.

# Exercise 6: Given a sample size of 60, 1000 bootstrap samples for each interval, and 50 confidence intervals constructed (the default values for the above app), what proportion of your confidence intervals include the true population proportion? Is this proportion exactly equal to the confidence level? If not, explain why. Make sure to include your plot in your answer.

** Ans6: 96% of my confidence intervals include the true population proportion. The proportion is not exactly equal to the confidence level.I wouldn’t expect that to be exactly equal to the confidence interval as because I was at least 95% certain that my confidence intervals would include the true population but it could be more than 95%. 

# Exercise 7: Choose a different confidence level than 95%. Would you expect a confidence interval at this level to me wider or narrower than the confidence interval you calculated at the 95% confidence level? Explain your reasoning.

** Ans7: I chose 88% confidence level and I would expect that a confidence interval at this level will be narrower than the confidence interval I calculated at the 95% confidence level. With 95% confidence level, there is a chance of 5% result to be wrong, whereas, with 88% confidence level, chances of result to be wrong is 12%. Hence, as the confidence level decreases the reliability of an interval containing the true population proportion also decreases.

# Exercise 8: Using code from the infer package and data from the one sample you have (samp), find a confidence interval for the proportion of US Adults who think climate change is affecting their local community with a confidence level of your choosing (other than 95%) and interpret it.

```{r }
set.seed(9087)

samp %>%
  specify(response = climate_change_affects, success = "Yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.88)

```

** Ans8: With 88% confidence level I have received a confidence interval, ci (0.52, 0.72). With this interval I am 88% confident that the true proportion of US adults who think climate change affects their local community will be contained in this interval.

# Exercise 9: Using the app, calculate 50 confidence intervals at the confidence level you chose in the previous question, and plot all intervals on one plot, and calculate the proportion of intervals that include the true population proportion. How does this percentage compare to the confidence level selected for the intervals?

** Ans9: By running the app with 88% confidence level, it is found that the percentage of intervals that include the true population proportion is lower (86%) than the confidence intervals with 95% confidence level (it was 96%)

# Exercise 10: Lastly, try one more (different) confidence level. First, state how you expect the width of this interval to compare to previous ones you calculated. Then, calculate the bounds of the interval using the infer package and data from samp and interpret it. Finally, use the app to generate many intervals and calculate the proportion of intervals that are capture the true population proportion.

```{r }
set.seed(9087)
samp %>%
  specify(response = climate_change_affects, success = "Yes") %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.98)

```
** Ans 10: With 98% confidence level, I will get wider interval as compared to the previous ones.From the confidence interval, ci(0.45, 0.77), it is clearly seen that this bound is more wider than the previuos bounds. I can say, that it is 98% certian that the true population proportion will be contained within this confidence interval. By ruuning tha app with a sample size of 60, 1000 bootstrap samples for each interval, and 50 confidence intervals it is found that 96% of the confidence interval captured the true proportion of population.

# Exercise 11: Using the app, experiment with different sample sizes and comment on how the widths of intervals change as sample size changes (increases and decreases).

** Ans 11: The sample size increases when the width of confidence intervals decreases. As the sample size decreases, the width of confidence intervals increases.

# Exercise 12: Finally, given a sample size (say, 60), how does the width of the interval change as you increase the number of bootstrap samples. Hint: Does changing the number of bootstap samples affect the standard error?

** Ans12: For a given sample size, increasing the number of bootstap will decrease the standard error.Hence, the sampling distributions will be narrow. That means the width of the interval will decrease.


